\section{Group Schemes}


\subsection{correspondence between Hopf algebras and affine group schemes}


We want to show that Hopf algebras correspond to affine group schemes.
This can easily be done categorically assuming both categories on either side are defined thoughtfully.
However, the categorical version will not be workable with if we do not also have links to the non-categorical notions.
Therefore, one solution would be to build the left, top and right edges of the following
diagram so that the bottom edge can be obtained by composing the three:
%   Cogrp Mod_R   ⥤        Grp Sch_{Spec R}
%       ↑ ↓                        ↓
% R-Hopf algebras → Affine group schemes over Spec R


\subsubsection{Bundling/unbundling Hopf algebras}


First, let's do the left edge.


\begin{proposition}[The antipode is a antihomomorphism]
  \label{0-hopf-antipode-antihom}
  \uses{}
  \lean{HopfAlgebra.antipode_mul_antidistrib, HopfAlgebra.antipode_mul_distrib}
  \leanok

  If $A$ is a $R$-Hopf algebra, then the antipode map $s : A \to A$ is anti-commutative, ie $s(a * b) = s(b) * s(a)$. If further $A$ is commutative, then $s(a * b) = s(a) * s(b)$.
\end{proposition}
\begin{proof}
  \uses{}
  % \leanok

  Any standard reference will have a proof.
\end{proof}


\begin{proposition}[Hopf algebras are cogroup objects in the category of algebras]
  \label{0-hopf-cogrp-alg}
  \uses{}
  \lean{instGrp_ClassOppositeCommAlgCarrierOpOf,
        instIsMon_HomOppositeCommAlgCarrierOpOfHomToAlgHomBialgHom,
        instHopfAlgebraCarrierCarrierUnopCommAlg,
        IsMon_Hom.toBialgHom}
  \leanok

  From a $R$-Hopf algebra, one can build a cogroup object in the category of $R$-algebras.

  From a cogroup object in the category of $R$-algebras, one can build a $R$-Hopf algebra.
\end{proposition}
\begin{proof}
  \uses{0-hopf-antipode-antihom}
  \leanok

  Turn the arrows around.
\end{proof}


\subsection{Spec of a Hopf algebra}


Second, let's do the top edge.


\begin{proposition}[Sliced adjoint functors]
  \label{0-slice-adj}
  \uses{}
  % \lean{}
  % \leanok

  If $a : F \vdash G$ is an adjunction between $F : C \to D$ and $G : D \to C$ and $X : C$, then there is an adjunction between $F / X : C / X \to D / F(X)$ and $G / X : D / F(X) \to C / X$.
\end{proposition}
\begin{proof}
  \uses{}
  % \leanok

  See https://ncatlab.org/nlab/show/sliced+adjoint+functors+--+section.
\end{proof}


\begin{proposition}[Limit-preserving functors lift to over categories]
  \label{0-over-lim}
  \uses{}
  \lean{CategoryTheory.Limits.PreservesLimitsOfShape.overPost, CategoryTheory.Limits.PreservesLimitsOfSize.overPost, CategoryTheory.Limits.PreservesFiniteProducts.overPost}
  \leanok

  If $F : C \to D$ is a functor preserving limits of shape $J$, then so is the obvious functor $C / X \to D / F(X)$.
\end{proposition}
\begin{proof}
  \uses{}
  % \leanok

  Hopefully easy.
\end{proof}


\begin{proposition}[Fully faithful product-preserving functors lift to group objects]
  \label{0-full-faithful-grp}
  \uses{}
  \lean{CategoryTheory.Functor.Faithful.mapGrp, CategoryTheory.Functor.Full.mapGrp}
  \leanok

  If a finite-products-preserving functor $F : C \to D$ is fully faithful, then so is $\Grp(F) : \Grp C \to \Grp D$.
\end{proposition}
\begin{proof}
  \uses{}
  \leanok

  Faithfulness is immediate.

  For fullness, assume $f : F(G) \to F(H)$ is a morphism. By fullness of $F$, find $g : G \to H$ such that $F(g) = f$. $g$ is a morphism because we can pull back each diagram from $D$ to $C$ along $F$ which is faithful.
\end{proof}


\begin{proposition}[$\Spec$ as a functor on algebras]
  \label{0-spec-alg}
  \uses{}
  \lean{algSpec, algSpec.instPreservesLimits, algSpec.instFull, algSpec.instFaithful, algSpec.fullyFaithful}
  \leanok

  $\Spec$ is a fully faithful contravariant functor from the category of $R$-algebras to the category of schemes over $\Spec_R$, preserving all limits.
\end{proposition}
\begin{proof}
  \uses{0-slice-adj, 0-over-lim}
  \leanok

  $\Spec : \Ring \to \Sch$ is a fully faithful contravariant functor which preserves all limits, hence so is $\Spec : \Ring_R \to \AffSch_{\Spec R}$ by Proposition \ref{0-over-lim} (alternatively, by Proposition \ref{0-slice-adj}).
\end{proof}


\begin{proposition}[$\Spec$ as a functor on Hopf algebras]
  \label{0-spec-hopf}
  \uses{}
  \lean{hopfSpec, hopfSpec.instFull, hopfSpec.instFaithful, hopfSpec.fullyFaithful}
  \leanok

  $\Spec$ is a fully faithful contravariant functor from the category of $R$-algebras to the category of group schemes over $\Spec_R$.
\end{proposition}
\begin{proof}
  \uses{0-full-faithful-grp, 0-spec-alg}
  \leanok

  $\Spec : \Ring_R \to \Sch_{\Spec R}$ is a fully faithful contravariant functor preserving all limits according to Proposition \ref{0-spec-alg}, therefore $\Spec : \Hopf_R \to \GrpSch_{\Spec R}$ too is fully faithful according to \ref{0-full-faithful-grp}.
\end{proof}


\subsection{Essential image of $\Spec$ on Hopf algebras}


Finally, let's do the right edge.


\begin{proposition}[Essential image of a sliced functor]
  \label{0-ess-image-over}
  \uses{}
  % \lean{}
  % \leanok

  If $F : C \to D$ is a fully faithful functor between cartesian-monoidal categories, then $F / X : C / X \hom D / F(X)$ has the same essential image as $F$.
\end{proposition}
\begin{proof}
  \uses{}
  % \leanok

  Transfer all diagrams.
\end{proof}


\begin{proposition}[Equivalences lift to group object categories]
  \label{0-grp-equiv}
  \uses{}
  \lean{CategoryTheory.Equivalence.mapGrp}
  \leanok

  If $e : C \backsimeq D$ is an equivalence of cartesian-monoidal categories, then $\Grp(e) : \Grp(C) \backsimeq \Grp(D)$ too is an equivalence of categories.
\end{proposition}
\begin{proof}
  \uses{}
  \leanok

  Transfer all diagrams.
\end{proof}


\begin{proposition}[Essential image of a functor on group objects]
  \label{0-ess-image-grp}
  \uses{}
  \lean{CategoryTheory.Functor.essImage_mapGrp}
  \leanok

  If $F : C \to D$ is a fully faithful functor between cartesian-monoidal categories, then $\Grp(F) : \Grp(C) \hom \Grp(D)$ has the same essential image as $F$.
\end{proposition}
\begin{proof}
  \uses{0-grp-equiv}
  % \leanok

  Transfer all diagrams.
\end{proof}


\begin{proposition}[Essential image of $\Spec$ on algebras]
  \label{0-ess-image-spec-alg}
  \uses{0-spec-alg}
  \lean{essImage_algSpec}
  \leanok

  The essential image of $\Spec : \Ring_R \to \Sch_{\Spec R}$ is precisely affine schemes over $\Spec R$.
\end{proposition}
\begin{proof}
  \uses{0-ess-image-over}
  % \leanok

  Direct consequence of Proposition \ref{0-ess-image-over}.
\end{proof}


\begin{proposition}[Essential image of $\Spec$ on Hopf algebras]
  \label{0-ess-image-spec-hopf}
  \uses{0-spec-hopf}
  \lean{essImage_hopfSpec}
  \leanok

  The essential image of $\Spec : \Hopf_R \to \GrpSch_{\Spec R}$ is precisely affine group schemes over $\Spec R$.
\end{proposition}
\begin{proof}
  \uses{0-ess-image-grp, 0-ess-image-spec-alg}
  \leanok

  Direct consequence of Propositions \ref{0-ess-image-grp} and \ref{0-ess-image-spec-alg}.
\end{proof}


\subsection{Diagonalisable groups}


\begin{definition}[Group-like elements]
  \label{0-grp-like}
  \uses{}
  \lean{Coalgebra.IsGroupLikeElem}
  \leanok

  An element $a$ of a coalgebra $A$ is \emph{group-like} if it is a unit and $\Delta(a) = a \ox a$, where $\Delta$ is the comultiplication map.
\end{definition}


\begin{lemma}[Bialgebra homs preserve group-like elements]
  \label{0-grp-like-map}
  \uses{0-grp-like}
  \lean{Coalgebra.IsGroupLikeElem.map}
  \leanok

  Let $f : A \to B$ be a bi-algebra hom. If $a \in A$ is group-like, then $f(a)$ is group-like too.
\end{lemma}
\begin{proof}
  \uses{}
  \leanok

  $a$ is a unit, so $f(a)$ is a unit too. Then
  \[
    f(a) \ox f(a) = (f \ox f)(\Delta_A(a)) = \Delta_B(f(a))
  \]
  so $f(a)$ is group-like.
\end{proof}


\begin{lemma}[Independence of group-like elements]
  \label{0-grp-like-lin-indep}
  \uses{0-grp-like}
  \lean{Coalgebra.linearIndepOn_isGroupLikeElem}
  \leanok

  The group-like elements in $A$ are linearly independent.
\end{lemma}
\begin{proof}
  \uses{0-tensor-lin-indep}
  \leanok

  See Lemma 4.23 in \cite{Milne_2017}.
\end{proof}


\begin{lemma}
  \label{0-grp-like-grp-alg}
  \uses{0-grp-like}
  \lean{MonoidAlgebra.isGroupLikeElem_iff_mem_range_of}
  \leanok

  The group-like elements of $k[M]$ are exactly the image of $M$.
\end{lemma}
\begin{proof}
  \uses{0-grp-like-lin-indep}
  \leanok

  See Lemma 12.4 in \cite{Milne_2017}.
\end{proof}


\begin{definition}
  \label{0-spec-grp-alg}
  \lean{AlgebraicGeometry.Spec, MonoidAlgebra}
  \leanok

  For a commutative group $M$ we define $D_R(M)$ as the spectrum $\Spec R[M]$ of the group algebra $R[M]$.
\end{definition}


\begin{definition}
  \label{0-diag}
  \uses{0-spec-grp-alg}
  \lean{AlgebraicGeometry.Scheme.IsDiagonalisable}
  % \leanok

  An algebraic group $G$ over $\Spec R$ is {\bf diagonalisable}
  if it is isomorphic to $D_R(M)$ for some commutative group $M$.
\end{definition}


\begin{theorem}
  \label{0-diag-iff-grp-like-span}
  \uses{0-spec-grp-alg, 0-diag}
  \lean{AlgebraicGeometry.Scheme.isDiagonalisable_iff_span_isGroupLikeElem_eq_top}
  \leanok

  An algebraic group $G$ over a field $k$ is diagonalizable if and only if group-like elements span $\Gamma(G)$.
\end{theorem}
\begin{proof}
  \uses{0-grp-like-lin-indep}
  % \leanok

  See Theorem 12.8 in \cite{Milne_2017}.
\end{proof}


\begin{proposition}[The group algebra functor is fully faithful]
  \label{0-full-grp-hopf-alg}
  \uses{}
  % \lean{}
  % \leanok

  The functor $G \rightsquigarrow R[G]$ from the category of groups to the category of Hopf algebras over $R$ is fully faithful.
\end{proposition}
\begin{proof}
  \uses{0-diag-iff-grp-like-span, 0-grp-like-grp-alg, 0-grp-like-map}
  % \leanok

\end{proof}


\begin{theorem}
  \label{0-fg-comm-grp-equiv-diag-grp-sch}
  \uses{1-1-char-lat}
  \uses{0-diag}
  The functor $M\rightsquigarrow D(M)$ is a contravariant equivalence
  from the category of finitely generated commutative groups to the category of
  diagonalizable algebraic groups (with quasi-inverse $G \rightsquigarrow X(G)$).
\end{theorem}
\begin{proof}
  \uses{0-hopf-aff-grp-sch, 0-full-grp-hopf-alg}
  % \leanok

  The functor is essentially surjective by definition.
  It's faithfull since, $M\rightsquigarrow k[M]$ and $\Spec$ are.
  It's full by \ref{congr_hopf_affgrpsch} and \ref{0-full-grp-hopf-alg}.

  Also, see Theorem 12.9(a) in \cite{Milne_2017}.
\end{proof}
