\section{Objects}


\subsection{Group objects}


\begin{proposition}[Fully faithful product-preserving functors lift to monoid/group objects]
  \label{0-full-faithful-grp}
  \uses{}
  \lean{CategoryTheory.Functor.Faithful.mapGrp, CategoryTheory.Functor.Full.mapGrp}
  \leanok

  If a finite-products-preserving functor $F : C \to D$ is fully faithful,
  then so is $\Grp(F) : \Grp C \to \Grp D$.
\end{proposition}
\begin{proof}
  \uses{}
  \leanok

  Faithfulness is immediate.

  For fullness, assume $f : F(G) \to F(H)$ is a morphism. By fullness of $F$,
  find $g : G \to H$ such that $F(g) = f$. $g$ is a morphism
  because we can pull back each diagram from $D$ to $C$ along $F$ which is faithful.
\end{proof}


\begin{proposition}[Equivalences lift to monoid/group object categories]
  \label{0-grp-equiv}
  \uses{}
  \lean{CategoryTheory.Equivalence.mapGrp}
  \leanok

  If $e : C \backsimeq D$ is an equivalence of cartesian-monoidal categories,
  then $\Grp(e) : \Grp(C) \backsimeq \Grp(D)$ too is an equivalence of categories.
\end{proposition}
\begin{proof}
  \uses{}
  \leanok

  Transfer all diagrams.
\end{proof}


\begin{proposition}[Essential image of a functor on monoid/group objects]
  \label{0-ess-image-grp}
  \uses{}
  \lean{CategoryTheory.Functor.essImage_mapGrp}
  \leanok

  If $F : C \to D$ is a fully faithful functor between cartesian-monoidal categories,
  then $\Grp(F) : \Grp(C) \hom \Grp(D)$ has the same essential image as $F$.
\end{proposition}
\begin{proof}
  \uses{0-grp-equiv}
  \leanok

  Transfer all diagrams.
\end{proof}


\begin{lemma}[A fully faithful functor product-preserving is a group isomorphism on hom sets]
  \label{0-full-faithful-grp-hom}
  \uses{0-full-faithful-grp}
  \lean{CategoryTheory.Functor.FullyFaithful.homMulEquiv}
  \leanok

  If $F : C \to D$ is a fully faithful functor between cartesian-monoidal categories
  and $X, G \in C$ are an object and a group object respectively,
  then $\Grp(F) : (X \hom G) \hom \backcong (F(X) \hom F(G))$ is a group isomorphism.
\end{lemma}
\begin{proof}
  \uses{}
  \leanok

  Toddlers and streets.
\end{proof}
